%!TEX root = ../report.tex

\documentclass[../report.tex]{subfiles}
\begin{document}

This Annual Progress Report, the first of our PhD programme, serves to detail the work we have done in the first year of our research. We first outline the research title we have worked under and outline the structure of the discussion to follow.

\subsection{Thesis Statement}

Our working research title is {\it game-theoretic and probabilistic methods applied to spatial network models of contagion.} Over the past year, we have spent time reading and developing ideas related to graph models of contagion, such as {\scshape Firefighter} and similar games on graphs. We have explored optimal strategies and probabilistic methods in order to adapt these games to stochastic contexts. Such a broad approach has meant that we now have a wide range of tools available to us in understanding and developing models of disease and other contagions.

\subsection{Report Roadmap}

Our research over the past year has focused into two main strands: a formal, mathematical approach and an experimental approach. Our motivation for this is to ultimately develop a model for disease spread formally, using dynamical system and graph theoretic approaches, which we then implement experimentally to explore different approaches to containment. One key benefit of this approach is that experimental data can be fed into a graph model of disease very quickly, contextualising the results of the model and informing the expected outcomes of approaches to containing the disease very rapidly, which is especially important in epidemic scenarios \cite{firestone_2019}. To explain the work we have done in detail, we will begin by discussing the literature examined in the past year, with particular focus on papers that have proved foundational, to provide a background to the work we have done in extending these results and concepts. We then discuss the work we have done in our research, including some mathematical results and some key experimental implementations and relevant results. We will then outline the direction we hope to take this work moving forward and detail milestones we hope to reach in the coming year.

\end{document}