%!TEX root = ../report.tex

\documentclass[../report.tex]{subfiles}
\begin{document}

\subsection{Goals}

Over the next year, we hope to extend the two strands of work we have been pursuing with a number of goals and milestones in mind. In the more formal, mathematical side of our work, we aim to achieve the following:
\begin{itemize}
	\item Formalise a probabilistic graph model that can be implemented in the PRISM model checking language.
	\item Complete and verify the equations generation work that we have started over the last year. 
	\item Write down the algorithmic approach used in generating required differential equations for Markovian compartmental epidemic graph models.
\end{itemize}

Further, in the experimental side of our work, we aim to achieve the following:
\begin{itemize}
	\item Further generalise the current implementation to accept further parameters and output requested results.
	\item Use the large computing capabilities of the FATA nodes to run many more models and analyse results.
	\item Implement, test and explore further defence strategies in the implementation, using this to reason about optimal defence candidates in different contexts.
	\item Write down a series of (provable) statements regarding defence strategies in this implementation on various graph classes.
	\item Use real-world data to model different epidemic scenarios and contextualise defence strategies.
	\item Explore integer programming approaches, such as provided in Finbow and MacGillivray \cite{finbow_2009}, and identify similar approaches for our extended model.
\end{itemize}


\subsection{Resources required}

\subsection{Measures of success}

\subsection{Timeline}

\end{document}